%% Appendix
%\appendix

\section{Proof: Bidirectional Optimizatoin Correctness}
\label{appendix:BidirectionalProof}





	\subsection{Preliminaries}
	
	\begin{definition}[Petri Net]
		A \emph{Petri net} is a tuple
		\[
		N = (P,\,T,\,\Pre,\,\Post)
		\]
		where
		\begin{itemize}
			\item $P$ is a finite set of \emph{places},
			\item $T$ is a finite set of \emph{transitions},
			\item $\Pre: P\times T \to \mathbb{N}$ is the \emph{pre-incidence} function,
			\item $\Post: P\times T \to \mathbb{N}$ is the \emph{post-incidence} function.
		\end{itemize}
	\end{definition}
	
	\begin{definition}[Marking]
		A \emph{marking} is a function $M: P \to \mathbb{N}$. We write $M(p)$
		for the number of tokens in place $p$.  The initial marking is
		denoted $M_0$.  A transition $t\in T$ is \emph{enabled} at marking
		$M$ if $\forall p\in P:\,M(p)\ge\Pre(p,t)$.  Firing $t$ yields the
		new marking
		\[
		M' = M - \Pre(\cdot,t) + \Post(\cdot,t),
		\]
		written $M \xrightarrow{t} M'$.
	\end{definition}
	
	\begin{definition}[Firing Sequence]
		A sequence $\sigma = t_1 t_2 \cdots t_k \in T^*$ is \emph{fireable}
		from $M_0$ if there exist markings $M_1,\dots,M_k$ such that
		$M_0\xrightarrow{t_1}M_1\cdots\xrightarrow{t_k}M_k$.  We write
		$M_0 \xrightarrow{\sigma} M_k$.
	\end{definition}
	
	\begin{definition}[Semilinear Target Set]
		A \emph{semilinear} set $T\subseteq \mathbb{N}^P$ is a finite union of
		linear sets.  We assume $T$ is given by a finite description of its
		linear components.  We view $T$ as the \emph{target} set of markings
		we wish to reach.
	\end{definition}
	
	\subsection{The Bidirectional Pruning Algorithm}
	
	Let $N=(P,T,\Pre,\Post)$, initial marking $M_0$, and target set
	$T\subseteq\mathbb{N}^P$ be fixed.
	
	\begin{definition}[Forward Over‐Approximation]
		Define the operator $\mathcal{F}:\mathcal{P}(P\cup T)\to\mathcal{P}(P\cup T)$ by
		\[
		X \mapsto X
		~\cup~
		\{\,t\in T \mid \forall p\in P:\; \Pre(p,t)>0 \implies p\in X\}
		~\cup~
		\{\,p\in P \mid \exists t\in X\cap T,\ \Post(p,t)>0\}.
		\]
		Starting from $X_0 = \{\,p\mid M_0(p)>0\}$, iterate
		$X_{i+1} = \mathcal{F}(X_i)$ until a least fixed‐point
		$X^*=\bigcup_i X_i$ is reached.  Call $X^*_P = X^*\cap P$ the set of
		\emph{forward‐reachable} places.
	\end{definition}
	
	\begin{definition}[Backward Over‐Approximation]
		Let
		\[
		Y_0 = \{\,p\in P \mid \exists M\in T:\;M(p)>0\}
		\]
		be the places unconstrained to zero by the target.  Define
		$\mathcal{B}:\mathcal{P}(P\cup T)\to\mathcal{P}(P\cup T)$ by
		\[
		Y \mapsto Y
		~\cup~
		\{\,t\in T \mid \forall p\in P:\; \Post(p,t)>0 \implies p\in Y\}
		~\cup~
		\{\,p\in P \mid \exists t\in Y\cap T,\ \Pre(p,t)>0\}.
		\]
		Iterate $Y_{i+1} = \mathcal{B}(Y_i)$ until a least fixed‐point
		$Y^*=\bigcup_i Y_i$ is reached.  Call $Y^*_P = Y^*\cap P$ the set of
		\emph{backward‐relevant} places.
	\end{definition}
	
	\begin{definition}[Pruned Net]
		The \emph{pruned} subnet is
		\[
		N' = \bigl(P',\,T',\,\Pre|_{P'\times T'},\,\Post|_{P'\times T'}\bigr)
		\]
		where
		\[
		P' = X^*_P \;\cap\; Y^*_P,
		\quad
		T' = \{\,t\in T \mid
		\forall p:\;\Pre(p,t)>0\implies p\in P',\;
		\forall p:\;\Post(p,t)>0\implies p\in P'
		\}.
		\]
	\end{definition}
	
	\subsection{Invariant and Correctness}
	
	Intuitively, $P'$ contains exactly those places that
	\emph{may} occur in some firing sequence from $M_0$ to a marking in $T$.
	
	\begin{definition}[Witnessable Place]
		A place $p\in P$ is \emph{witnessable} if there exists a firing
		sequence $\sigma\in T^*$ and markings $M$ and $M'$ such that
		\[
		M_0 \xrightarrow{\sigma_1} M
		\quad\text{and}\quad
		M \xrightarrow{\sigma_2} M'
		\quad\text{with}\quad
		M'(p)>0
		\quad\text{and}\quad
		M'\in T.
		\]
		In other words, $p$ can carry a token in some execution from $M_0$ into the target set.
	\end{definition}
	
	\begin{theorem}[Pruning Invariant]
		\label{thm:invariant}
		If a place $p$ is witnessable, then $p\in P'$.  Equivalently, the
		pruned net $N'$ \emph{over‐approximates} the set of witnessable places.
	\end{theorem}
	
	\begin{proof}
		We split the argument into two parts.
		
		\paragraph{(1) Forward‐reachability.}
		Suppose $p$ is witnessable.  Then there is a prefix
		$\sigma_1\in T^*$ such that $M_0\xrightarrow{\sigma_1}M$ and
		$M(p)>0$.  By standard Petri‐net monotonicity, every place that
		receives a token in the course of $\sigma_1$ must appear in the
		forward fixed‐point $X^*_P$.  Hence $p\in X^*_P$.
		
		\paragraph{(2) Backward‐relevance.}
		Again, since $p$ is witnessable, there is a suffix
		$\sigma_2\in T^*$ from $M$ to $M'\in T$ with $M'(p)>0$.  Working
		backwards from $T$, every place that can contribute to satisfying
		semilinear constraints (i.e.\ those places that might be tested
		positive in $T$) appears in the backward fixed‐point $Y^*_P$.  Thus
		$p\in Y^*_P$.
		
		\paragraph{Conclusion.}
		Combining (1) and (2) yields $p\in X^*_P\cap Y^*_P = P'$, as desired.
	\end{proof}
	
	\subsection{Termination and Complexity}
	
	\begin{lemma}
		Each iteration of $\mathcal{F}$ and $\mathcal{B}$ strictly increases
		the set of included elements (unless already at the fixed‐point), and
		the total number of elements is finite.  Hence both reach their
		fixed‐points in at most $|P|+|T|$ iterations each.
	\end{lemma}
	
	\begin{proof}
		Immediate from monotonicity and finiteness.
	\end{proof}
	
	\noindent
	Therefore the bidirectional pruning converges in polynomial time
	(mostly linear in the net‐size per pass), and preserves exactly the
	places and transitions that \emph{may} appear in some execution from
	$M_0$ into $T$.
	

