% ---- Document Class ----
\documentclass[sn-mathphys,Numbered]{sn-jnl}

% ---- Essential Packages ----
% FIX: Load xcolor ONCE with all necessary options to prevent "Option Clash"
\usepackage[table,dvipsnames,svgnames]{xcolor}

\usepackage{amsmath}
% FIX: Handle \Bbbk conflict before loading amssymb
\let\Bbbk\relax 
\usepackage{amssymb}
\usepackage{mathtools}
\usepackage{amsthm}

\usepackage[demo]{graphicx} % 'demo' allows compiling without images
\usepackage{hyperref}

% ---- Formatting & Layout ----
\usepackage{alltt}
\usepackage{paralist}
\usepackage{booktabs}
\usepackage{multirow}
\usepackage{arydshln}
\usepackage{wrapfig}
\usepackage[export]{adjustbox}
\usepackage{enumitem}
\usepackage{float}
\usepackage{caption}
\usepackage{subcaption}
\usepackage{tcolorbox}
\usepackage{nicefrac}

% ---- Code Listings ----
\usepackage{listings}
\usepackage{verbatim}

% ---- Math & Symbols ----
\usepackage{mathpartir}
\usepackage{wasysym}
\usepackage{pifont}
\usepackage{stmaryrd}
\usepackage{bussproofs}
\usepackage{gensymb}
\usepackage[per-mode=symbol,detect-all]{siunitx}

% ---- Other Utilities ----
\usepackage{xspace}
\usepackage{url}
\usepackage{svg}
\usepackage{ifthen}
\usepackage{etoolbox}
\usepackage[outline]{contour}

% ---- TikZ Graphics ----
\usepackage{tikz}
\usetikzlibrary{shapes.geometric,positioning,arrows}

% ---- Cleveref (Must be loaded after hyperref) ----
\usepackage[capitalize]{cleveref}

% ---- Custom Definitions & Macros ----

% Theorem environments
\newtheorem{theorem}{Theorem}[section]
\newtheorem{lemma}[theorem]{Lemma}
\newtheorem{proposition}[theorem]{Proposition}
\newtheorem{corollary}[theorem]{Corollary}
\theoremstyle{definition}
\newtheorem{definition}[theorem]{Definition}
\theoremstyle{remark}
\newtheorem{remark}[theorem]{Remark}

% Colors
\definecolor{brightyellow}{RGB}{255,247,192}
\definecolor{lightchacki}{RGB}{245,237,205}
\definecolor{darkorange}{RGB}{255,140,0}
\definecolor{headerbg}{RGB}{220,230,241}

% Symbols
\newcommand{\cmark}{\checkmark}
\newcommand{\xmark}{\textcolor{red}{\ding{55}}}
\newcommand{\greencmark}{\textcolor{green}{\ding{51}}}

% Math Macros
\newcommand{\cfg}[3]{\langle #1,\; #2,\; #3\rangle}
\newcommand{\update}[3]{#1[#2 \mapsto #3]}
\newcommand{\step}{\mathrel{\rightarrow}}
\newcommand{\pstep}{\mathrel{\Rightarrow}}
\newcommand{\Pre}{\mathsf{pre}}
\newcommand{\Post}{\mathsf{post}}
\newcommand{\toolname}{\textsc{Ser}}
\newcommand{\grammartag}[1]{\qquad\qquad\emph{(#1)}}

% Keywords
\newcommand{\kw}[1]{\textbf{#1}}
\newcommand{\nondet}{\kw{?}}
\newcommand{\ifkw}{\kw{if}}
\newcommand{\elsekw}{\kw{else}}
\newcommand{\whilekw}{\kw{while}}
\newcommand{\yieldkw}{\kw{yield}}
\newcommand{\requestkw}{\kw{request}}
\newcommand{\sat}{\texttt{SAT}}
\newcommand{\unsat}{\texttt{UNSAT}}
\newcommand{\Parikh}{\mathsf{Parikh}}

% Helper commands
\newcommand\blfootnote[1]{%
	\begingroup
	\renewcommand\thefootnote{}\footnote{#1}%
	\addtocounter{footnote}{-1}%
	\endgroup
}

\newcommand{\guy}[1]{\marginpar{\textcolor{orange}{Guy: #1}}}
\newcommand{\jules}[1]{\marginpar{\textcolor{red}{Jules: #1}}}
\newcommand{\markb}[1]{\marginpar{\textcolor{black}{Mark: #1}}}
\newcommand{\todo}[1]{\textcolor{red}{TODO: #1}}

% Listings Configuration
\lstdefinelanguage{CustomPseudoCode}{
	morekeywords={request, yield, return, if, else, while, and, or},
	morecomment=[l]{//},
	morestring=[b]",
	sensitive=true
}

\lstset{
	language=CustomPseudoCode,
	basicstyle=\ttfamily\small,
	keywordstyle=\color{blue}\bfseries,
	commentstyle=\color{gray}\itshape,
	stringstyle=\color{orange},
	numbers=left,
	numberstyle=\tiny,
	stepnumber=1,
	numbersep=5pt,
	backgroundcolor=\color{white},
	frame=single,
	rulecolor=\color{black},
	tabsize=2,
	captionpos=b,
	breaklines=true,
	breakatwhitespace=false,
	showstringspaces=false,
	escapeinside={(*@}{@*)},
}

% Hyperref fixes
\makeatletter
\renewcommand*{\theHsection}{\thesection}
\renewcommand*{\theHsubsection}{\thesubsection}
\renewcommand*{\theHsubsubsection}{\thesubsubsection}
\patchcmd{\@sect}
{\@tempskipa #5\relax}
{\refstepcounter{#1}%
	\Hy@raisedlink{\hyper@anchorstart{#1.\csname the#1\endcsname}\hyper@anchorend}%
	\@tempskipa #5\relax}
{}{}
\makeatother

% ==========================================
%            DOCUMENT START
% ==========================================
\begin{document}
	
	\title{Deciding Serializability in Network Systems}
	
	\author*[1]{\fnm{Guy}\sur{Amir}}\email{gda42@cornell.edu}
	\author[1]{\fnm{Mark}\sur{Barbone}}\email{mlb494@cornell.edu}
	\author[2]{\fnm{Nicolas}\sur{Amat}}\email{nicolas.amat@onera.fr}
	\author[1]{\fnm{Jules}\sur{Jacobs}}\email{jj758@cornell.edu}
	
	\affil[1]{\orgname{Cornell University}, \orgaddress{\city{Ithaca}, \country{United States}}}
	\affil[2]{\orgname{DTIS, ONERA, Universit\'e de Toulouse}, \orgaddress{\city{Toulouse}, \country{France}}}
	
	\abstract{
		We present the \toolname{} modeling language for automatically verifying \textit{serializability} of concurrent programs.
		\toolname{} programs are suitably restricted to make this problem decidable, while still allowing for an \textit{unbounded} number of concurrent threads.
		
	}
	
	\keywords{Serializability, Concurrent programs, Network systems, Petri nets, Decision procedure}
	
	\maketitle
	\pagestyle{plain}
	\thispagestyle{plain}
	
	% PLACEHOLDERS FOR CONTENT
	% (Uncomment your \input lines when running locally)
	
	\section{Introduction}
	% \input{sections/1_introduction}
	This is placeholder text for the introduction.
	
	\section{Problem Definition}
	% \input{sections/2_problem_definition}
	This is placeholder text for the problem definition.
	
	
	\section{Formal Results}
	% \input{sections/3_formal_results}
	This is placeholder text for formal results.
	
	\section{Implementation}
	% \input{sections/4_implementation}
	This is placeholder text for implementation details.
	
	\section{Evaluation}
	% \input{sections/5_evaluation}
	This is placeholder text for evaluation.
	
	\section{Related Work}
	% \input{sections/6_related_work}
	This is placeholder text for related work.
	
	\section{Discussion}
	% \input{sections/7_discussion}
	This is placeholder text for discussion.
	
	\section*{Data and Software Availability}
	The data and software are available.
	
	\section*{Acknowledgements}
	The work of Amir was partially supported by a Rothschild Fellowship.
	
	% Bibliography
	% \bibliographystyle{splncs04}
	% \bibliography{references}
	
	\newpage
	\appendix
	\section{Appendix}
	This is placeholder text for the appendix.
	
\end{document}